\subsection{Alternative Implementations}

\subsubsection{Bootstrapping off of Existing Storage Providers}

It is possible to implement Graffiti on top of existing storage providers,
like Dropbox or Solid. When a user posts an object, that object is stored in
a large folder of all the user's objects. Then that object is symlinked into
folders for each channel the user wants to post the object to.

Unlike in our canonical implementation, it is not possible to query for
all objects in a channel by a particular storage provider. Instead, a
separate BitTorrent-like \emph{tracker} service is required to
point users to the appropriate storage provider channel.

Using a commercial storage provider like Dropbox allows for access control but only if users
have a Dropbox account, which locks users into a proprietary service.
Solid does not have this problem. In fact, it is not possible to use the Dropbox API
without being used into a Dropbox account, but it is possible to use the Solid API
without being locked into a Solid account.

The primary reason for not pursuing this implementation is that it requires
an extremely large number of network requests to the storage providers
that is both expensive and slow. Additionally, these storage providers
do not support querying for objects in a particular channel and instead
all the filtering has to be done client-side.

While we expect some users of our canonical implementation to host their own pods,
many will use large pods hosted by others, much like Email or Mastodon.
Only a single network request is required per pod to retrieve all objects in a channel,
whereas a seperate request is required per user and event per object in the Dropbox model.

\subsubsection{Peer-to-Peer Implementation}

While this may have been possible in the PC era, it is not possible in the mobile era.
Most users are not online all the time and can lose their devices or accidentally wipe their data.
