\subsection{Core Client Library}

In addition to interfacing with identity providers,
and storage pods, the core library provides additional functionality
around delegation, announcements, filtering and caching.

The library is written in Typescript and can be used in both the browser and node.js.

\subsubsection{Delegation}

Many decentralized systems use public key cryptography to sign messages,
however this can mean that anything a user writes is undeniable and can
possibly be leaked even if it was meant to be private as we discussed in X.
Instead we use delegation, whereby a user chooses.

Therefore, the client application needs to know which pods a user has delegated
to so that they can know whether to trust the data.

As we saw in \cref{identity-provider}, the Solid OIDC specification
allows a user to store some amount of data at their webId.
That data is a URL which points to a Graffiti object owned by the user
on some pod, and includes global settings.
These settings include a list of

The settings also include forwarding rules which allow a user to

The verification of these settings is done by the client core library.

\subsubsection{Announcements}

When an application makes a \texttt{discover} request to a pod, that request
only returns results from that specific pod. However, users may have posted across
multiple pods. Therefore when a user makes a post in one pod, the client library
also puts "announcements" in several well known pods of the users choosing.
This way, when a user makes a \texttt{discover} request, their client first
makes \texttt{discover} requests for announcements in the users known pods
and then makes \texttt{discover} requests from the pods that have been found.
So long as there is some overlap in the well known pods that users have chosen,
the user will be able to see all of their posts.

These announcements are graffiti objects like any other and so more metadata
or announcement-like objects can be added over time to improve efficiency or
discoverability. For example, it may not be necessary.

\subsubsection{Filtering}

Types help to provide a better software experience.
Anyone can host a pod and so the results are not necessarily trustworthy.
So

\subsubsection{Caching}

A users profile can be
