Graffiti is developed as an API which allows the underlying implementation
to evolve and change without breaking the applications.
Additionally, the existence of multiple implementations is proof that
the system model has been successfully abstracted away from the API.
This is not the case in, say, ActivityPub, where ActivityPub application developers
and even users need to be highly concious of the system model.
What servers their content lives on, who they are federating with, etc.

The implementations include:
- local
- monolithic
- commodity storage and tracker
- federated
- distributed

These implementations sketch out the design space and offer a,
described in the table below.

## Local

The local implementation is an implementation of Graffiti where all data is stored
locally and only availble on one device. While this seems strange for a tool
proported to build social applications, it is paramount to the developement experience.
It allows for rapid iteration and testing, and creation of users.

When a user calls "login()" in the local implementation, a popup appears that
simply asks them to type in their actor ID, no additional authentication is required.
They don't need to worry about polluting their own online presence with test data.
Once an application it can be swapped out seamlessly with any of the other applications
by just changing one line of code.

Our local implementation builds on PouchDB which provides persistent
storage both in the browser and in node.js.

A local implementation is possible to be used

Additionally, it also makes it possible to have hybrid implementations,
that draw from multiple sources.

## Monolithic

A traditional.
This may be useful in cases like Smalltown that want to host a social application but
intentionally want to "defederate" or be isolated from the network.

While our is not implemented, it is possible for monoliths to implement end-to-end encryption
However that may require the introduction of of public key infrastructure
which has usability problems discussed in distributed.

## Commodity Storage and Tracker

All data is stored in a commodity storage provider like Google Drive or Dropbox.
A tracker, like a bittorrent tracker, is

## Federated

This is federation of the monolithic storage providers.
Each federated pod is essentially an instance of a monolithic provider.
The only difference is that identity is portable.
In other words with one identity, you can get access controlled (see allowed) content
from any pod.

Like the commodity storage, it is still necessary to know which pods to query,
but unlike commodity storage an external tracker service is not necessary as it
can be built *on top* of the existing infrastructure.

## Distributed

This implementation is hypotehtical at the moment.


### Discussion

#### Economics

No metrics.
