A number of systems provide

\begin{table}[h]
    \centering
    \begin{tabular}{lccccc}
        \toprule
        \textbf{System} & \textbf{Accessible Design} & \textbf{Adversarial Interoperability} & \textbf{Context Differentiation} & \textbf{Autonomous Extensibility} & \textbf{Serverless} \\
        \midrule
        Firebase     & ● & ○ & ◐ & ○ & ● \\
        Deja Vue     & ● & ○ & ◐ & ○ & ◐ \\
        ActivityPub  & ○ & ○ & ◐ & ○ & ◐ \\
        Matrix       & ○ & ○ & ● & ○ & ◐ \\
        At Protocol  & ● & ● & ○ & ○ & ◐ \\
        Nostr        & ○ & ● & ○ & ○ & ◐ \\
        \bottomrule
    \end{tabular}
    \caption{Comparison of properties across different systems}
    \label{tab:systems_properties}
\end{table}

This differs from ActivityPub, where actors are associated with a particular `instance.'

One type of evolution that is required is evolution in data schemas.
This evolution rules out designs that have a fixed set of features, like diaspora*, or even designs that allow users to compose applications out of a pre-defined set of concepts, like
Déjà vu.
In either case, these functionalities can't be easily expanded. For example the `scoring' concept in Déjà Vu couldn't be expanded to include reactions without changing the core implementation of the concept itself or creating a new, and therefore backwards-incompatible concept.

\subsection{Pluralistic Infrastructures for Social Applications}

An \emph{infrastructure} for social applications
is the

to be the technologies upon which some set of social applications can be built.
Additionally, we expect applications built upon an infrastructure to interoperate easily, as opposed to applications built with a \emph{toolkit}, which we will review below.  A good infrastructure will offload significant effort for a developer but (somewhat in tension with this) will minimize the constraints the infrastructure imposes on the design of applications using it.  The IP routing infrastructure, for example, generally permits developers to ignore the details of packet delivery, but imposes no restrictions on what goes into a packet besides a size limit and a source and destination address.

Many of the infrastructures for major social applications are \emph{monolithic}, supporting only a single end-user application: there is only one experience for Facebook or TikTok.
Some infrastructures are quasi-monolithic and permit a narrow design space of social applications to be built upon them, such as Twitter and Reddit, which (before 2023) allowed third party apps to be built using their API.
A \emph{pluralistic} infrastructure is one that permits a broad design space of interoperable applications\footnote{The idea of pluralism comes from the indigenous anarchist revolutionaries, the Zapatistas. They capture pluralism in the phrase ``A world where many worlds fit''~\cite{designsforthepluriverse,firstworldhahaha}}.

% The ones we are interested in are pluralistic.
% Our goal for Graffiti is for it to be highly
% protocols, servers and other resources on top of which other social applications can be built.
% Reddit API and Twitter API.

In the following subsections we review infrastructures that demonstrate high degrees of pluralism.
It happens that all of these infrastructures are motivated in part, if not in full, by a distrust of \emph{centralized} social media infrastructures.
Centralized infrastructures are ones whose technology is controlled by a single entity, like a corporation.
This distrust is understandable as a malicious central entity
can interfere with content distribution, mine user data to develop algorithms for manipulation,
sell user data to third parties, or make themselves more monolithic as both Reddit and Twitter did in 2023 by limiting API access~\cite{ageofsurveillancecapitalism, platformsociety, standoutofourlight}.
%Not to mention handing data to third parties like banks or governments.
But while we hare this distrust, and consider it in our implementation, it is diversity rather than distrust that is the central motivator for this work.
% However, distrust is not the central adversary in our work.

Instead, we are concerned that infrastructures, whether centralized or decentralized, can be difficult to build upon and lock people in to certain design patterns --- in other words, we are concerned that existing infrastructures are not pluralistic enough.
Decentralized infrastructures can limit themselves to design patterns copied from the centralized services they were designed to replace or use technology for decentralization that introduces new limits on the user experience.
% For example, many of these systems have locked in designs that make user interactions highly public
% and many have introduced cryptography that makes posts undeniable and unforgettable.


% In the space of alternatike social media ecosystems, a common design goal is to uphold the security, privacy, and ownership of personal data.
% While we generally stand by these goals and consider them in the implementation of our system, they are orthogonal to primary goal of this work: making social applications easier to create, change, and migrate between.

% Concerns over security, privacy, and data `ownership stem from a distrust in platforms.
% For the sake of unraveling these motivations,
% we are not directlt concerned that Untrusted platforms may
% use user data to personalize user experiences and keep users on platforms longer, often with harmful externalities;
% they may give or sell data to third parties resulting in higher insurance premiums, rejection from jobs, schools, or loans, and incarceration;
% they may use user data to manipulate what users consume, how users spend money, who they meet, and what their public opinions are.

For example, two early decentralized infrastructures that we consider quasi-monolithic are diaspora* and Secure ScuttleButt.
These infrastructures bake in Twitter-like microblogging design patterns: a feed of content posted by people a user follows and interactions via comments and likes.
Both further limit the experience by signing all posts cryptographically with a user's public key. This makes all of a user's messages undeniable, unlike many centralized platforms where users can claim, for example, that a screenshot of a deleted post has been doctored.
ScuttleButt takes this a step further by using an append-only structure that makes it impossible to truly edit or delete content~\cite{diaspora, securescuttlebutt}.

% However, distrust is not the central adversary in our work. Instead, we are concerned that social infrastructures, regardless of their trustworthyness, can be difficult to build upon and lock people in to certain design patterns.
% Trustworthy services can limit themselves to design patterns copied from the distrustful services they were designed to replace or use technology to establish trust that introduces new limits on the user experience.
% For example, many of these systems have locked in high levels of content publicity and many have introduced cryptography that makes posts undeniable and unforgettable.

% Still, almost all existing systems that allow for more user security, privacy, and data ownership \emph{also} introduce more flexibility in developing end-user experiences than their centralized counterparts.
% We explore these systems in the following subsections.

% In our exploration of existing systems we will
So while all the infrastructures below are decentralized,
we will not touch touch on the mechanisms that make them decentralized except where those mechanisms limit or expand the end-user experience in some way.
We focus on what these infrastructures have done to expand the scope of possible application designs, and where they have room to grow that scope more.

% Such mechanisms include peer-to-peer interactions, federation, and end-to-end encryption.
% These mechanisms include peer-to-peer systems that do not rely on a central trusted authority; federated systems that are composed of many smaller authorities, one of which is a user can put trust in; and end-to-end encrypted systems that only reveal information to communicating parties, removing the need for trust in a platform.

% \subsubsection{diaspora*}

% Founded in 2010, diaspora* is one of the first distributed social networking protocols.
% diaspora* is a microblogging service like Twitter and includes concepts like profiles, posts, comments, and likes.
% diaspora* allows for content to shared either publicly, or to an `aspect': a reusable, asymmetric, and hidden set of users specified by the sender, supposedly the inspiration for the `circles' concept in Google+.
% % The stream of content a user receives is typically presented in a feed.

% Like many systems to follow, diaspora* signs all user interactions with the user's public key.
% While this is useful to prove authenticity in the distributed implimentation, it does mean that all of the user's messages are undeniable, unlike traditional social media platforms where users can claim, for example, that a screenshot of a deleted post has been doctored.

% The diaspora* protocol only defines server to server interactions rather than client to server interactions. This means that that application developers most likely need to roll their own servers to build out custom clients.

%https://wiki.diasporafoundation.org/FAQ_for_users
%https://diaspora.github.io/diaspora_federation/

% \subsubsection{Secure ScuttleButt}

% Founded in 2014, Secure ScuttleButt, offers a relatively similar experience to to diaspora* and Twitter with affordances to
% follow, post, reply, and like.
% Users mostly see content posted by people that they follow but there are built-in heuristics for getting content of `transitive interest' by friends of friends~\cite{securescuttlebutt}.

% ScuttleButt's unique peer-to-peer infrastructure gives it some limitations.
% First, ScuttleButt is an append only system.
% While some clients allow for users to edit or delete posts, the original posts are still in the system and can be found and attributed to the original user by anyone curious.
% Second, ScuttleButt requires knowing someone already in the network to bootstrap it's peer-to-peer network and begin receiving posts.
% Third, its architecture does not allow for central moderation, so moderation done by user flagging of harmful content.

% Like, diaspora*, ScuttleButt only specifies server-to-server interactions, requiring full stack development to create custom clients.

\subsubsection{ActivityPub}

Published as a World Wide Web Consortium recommendation in 2018, ActivityPub is the standard underlying the fediverse. The fediverse is an open ecosystem of applications including Mastodon (Twitter-like), Lemmy (Reddit-like), PeerTube (Youtube-like) and PixelFed (Instagram-like).
Meta's Threads is also rolling out fediverse support.
%https://engineering.fb.com/2024/03/21/networking-traffic/threads-has-entered-the-fediverse/

ActivityPub builds on its sister standards Activity Streams and Activity Vocabulary, schemas which provide a diverse and extensible set of \emph{objects} that can be created and \emph{activities} that can be performed by \emph{actors}, which may be users or other entities like communities.
Activities include Announce, Like or Follow, and objects include Note, Image, or Profile.
Activities and objects are represented in JSON-LD and each have standard associated properties like `content,' `tag,' and `inReplyTo.'
The `LD' in JSON-LD stands for `linked data' and allows for this schema to be extended and so
Mastodon, for example, has expanded the schema to include an Emoji object for posting emojis with custom icons and a `featured' property for pinning posts to the top of one's profile.
% https://docs.joinmastodon.org/spec/activitypub/

To share these activities and objects around, ActivityPub gives each actor an inbox for receiving messages and an outbox for broadcasting messages.
This model naturally allows for direct messaging: to message another user you put a message in their inbox, like an email; and friend/follower interactions: to subscribe to posts from another user you poll that user's outbox, like an RSS feed.
However, there are other interactions that cannot be easily simulated with inboxes and outboxes alone and so ActivityPub adds \emph{side effects} to activities and objects to compensate.

%This system gives ActivityPub a lot of flexibility, however the protocol is vague or silent on how many objects, activities, and properties should interpreted.
For example, a reply can be represented in the Activity Vocabulary as an Announce activity of a Note object that has an `inReplyTo' property pointing to the parent post that is being replied to.
However, if a user simply posts the Announce activity publicly to their outbox, how are viewers of the parent post supposed to know to look in the replying user's outbox?
The \emph{suggested} solution is that when a user posts a reply, it has the side effect of also being sent directly to the parent poster's inbox. When that parent poster receives the reply, they have the side effect of rebroadcasting it to their followers.
This generally works but number of edge cases can lead to `ghost replies' that implementers have built various heuristics to catch.
% https://seb.jambor.dev/posts/understanding-activitypub-part-3-the-state-of-mastodon/

%https://fedidb.org/software
Many ActivityPub activities have required or suggested side effects and additional side effects are added to make certain application-specific features possible.
Therefore there is no general purpose ActivityPub server.
We observe that the top 5 open source ActivityPub communities, according to monthly active users, Mastodon, Lemmy, PixelFed, Peertube, and Pleroma, all use their own custom server implementations, consisting of hundreds of thousands of lines of code each. Each implementation has custom side effects for application-specific message forwarding, moderation, content discovery, and aggregation.

Some of these applications, like Mastodon, have client-server APIs that allow users to build varied Mastodon clients, however like Twitter (before 2023), diaspora*, or Secure ScuttleButt, Mastodon on its own is quasi-monolithic.
Unlike diaspora* and Secure ScuttleButt, ActivityPub does not require messages to be signed with a public key, however most implementations do in fact mandate it.  This imposes constraints on client applications.   If an infrastructure supports anonymous messaging, it is easy to build clients that do authentication end-to-end using their own signatures; however, building signing into the infrastructure makes anonymous or deniable messaging impossible.

\subsubsection{Matrix}

Matrix was developed in 2014 as a real-time chat protocol, similar to IRC and XMPP.
It is so flexible however, that in 2020 the Matrix.org Foundation released a prototype microblogging application built upon the protocol.

Matrix is centered around the concept of `rooms'.
Rooms can be private and require invitation to enter or can be open to the public.
Messages shared to a particular room are shared with all of the other participants in the room.

The prototype microblogging app Cerulean works as follows:
each user creates a personal `timeline' room that is public but read-only to other users. Users publish their posts to their timeline room, similar to an ActivityPub outbox. One user can follow another by subscribing to messages from their timeline room.
Whenever someone creates a post, they must also create a public `thread' room for that post, for users to write comments in.

Matrix has a standardized API of creating rooms and forwarding messages between participants within in rooms, so Cerulean was implemented with frontend code over a \emph{generic} Matrix server.
However, the prototype is fairly limited and does not explain, for example how other features will work, like user discovery.

Additionally, Matrix is limited by a coarse moderation system.
Users within a room each have a `power level' from 0 to 100.
The creator of a room starts with a power level 100 and others start at level 0.
Different room actions can be assigned to require different power levels. By default, posting and inviting users to the room require power level 0, redacting messages or kicking users out of the room requires power level 50, and changing a person's or action's power level requires power level 100.
While this system may be enough for most chat applications, it is overly simplified for the complex moderation needs of large social networks~\cite{policykit}.   And because it is built in to the server, it is difficult if not impossible to design alternative moderation schemes for custom client applications.

% The Dutch
% PubHubs are built on top of Matrix. Homeservers serve their own clients. It uses it's own identity system.~\cite{pubhubs, pubhubsidentity}

As of 2020 Matrix uses public key encryption to sign all messages.

%https://matrix.org/blog/2020/12/18/introducing-cerulean/

\subsubsection{At Protocol}
https://docs.bsky.app/docs

Released in 2022, the At Protocol is the protocol underlying the Bluesky social network.
All posts in the At Protocol are public --- as of March 2024, there is not even private messaging.

All of a user's posts are kept in a personal data store that tracks changes like a git repository. All of these changes are signed with the user's public key which means posts are undeniable and unforgettable, since all deletions and edits a user makes are publicly available in this log.

Independent developers can build relays that index the content stored in individual data repositories to compute algorithmic feeds. Clients can be built on top of a relay's API by calling methods like `getTimeline` or `getPostThread.' However, relays don't necessarily need to share an API so there is no universal client-server API.
The relay-specific client-server APIs are at least be described by a common At Protocol-specific lexicon.

Relays may perform some amount of baseline moderation, but there is standardized labeling system that allows users to mark posts or other users as `rude,' `troll,' etc.
Users can subscribe to other users' labeling, allowing relays to filter or blur posts labeled by those users.
They call this system `composable moderation.'

\subsubsection{Nostr}

Released in 2020, Nostr's infrastructure, like the At Protocol's, is divided into three layers: data storage, relays, and clients.
Unlike the At Protocol, Nostr relays use a universal client-server API that allow users to query the relays for content they've indexed like a general purpose database.

For example, when one user follows another, their client simply queries the relays it is connected to for posts by that user.
This system is so generic it has allowed for at 78 different applications to be built on top of Nostr, including microblogging sites, Reddit-like forums, Pinterest-like collaborative tagging, an app for collaborative music remixing, and more.
%https://www.nostrapps.com/

However the flexibility to make arbitrary queries is also Nostr's drawback:
all of a user's activity across all client applications can be revealed by simply querying for their user ID.
Nostr provides basic mechanism for crafting private messages (which, by necessity, leaks a lot of metadata so the messages appear in queries)
but this won't prevent, for example, a user's posts in in a public political forum from mixing with their public life updates from mixing with their public dating profile --- this is the phenomenon of context collapse, which we will return to in \S.

Many of Nostr's users are cryptocurrency enthusiasts and cryptocurrency micropayments are a built-in feature of the platform. Unsuprisingly, all Nostr data is signed by a user's public key.
Similar to the At Protocol, users can opt-in to moderation by subscribing to block lists.

%https://github.com/nostr-protocol/nips/blob/master/45.md
%https://gregwhite.blog/nostr-content-moderation/

% \subsubsection{Spritely}

% Spritely was founded in 2020 by one of the cofounders of ActivityPub, although as of writing, there is no working implimentation.
% In an announcement, the founder shares similar concerns to our own:
% A co-author of the ActivityPub protocol wrote, ``contemporary fediverse interfaces borrow from surveillance-capitalism based popular social networks by focusing on breadth of relationships rather than depth.''
% For example, a co-author of the ActivityPub protocol later critiqued it, writing: ``Much of the fediverse has imported `what works' directly from Facebook and Twitter, but I’d argue there’s a lot we can do if we drop the assumption that this is the ideal starting base''~\cite{spritely}.

% Spritely uses object capabilities and is concerned with objects acting upon other objects in a globally consistent and permissioned way. This is important, for example, in social gaming whlre you would like to trade
% This could potentially be used to create actors, that have well defined behaviouhs (e.g. routing messages), but

% To our knowledge, spritely implimenters must run their own servers, cryptography

% https://spritely.institute/static/papers/spritely-core.html

\subsection{Social Application Frameworks}

There are a number of existing tools that allow users to more easily create personal social applications with little to no interoperation.
We call these tools \emph{frameworks} rather than infrastructures.
These frameworks are useful for building self-contained communities but are limited in their ability to fork into different designs without breaking their existing community apart.
% Many of these frameworks tend lo focus on low or no code development, a practice we intend to carry over more into infrastructure design.

Classic commercial tools like WordPress allow users to create sites with basic comment sections. With the help of plugins, like BuddyPress, users can add on pre-built social components like profiles, posting, and private messages.

The academic prototype Déjà Vu provides a catalog of primitive \emph{concepts} that offer more flexibility than components because they can be composed with each other to create complex high-level interactions. For example a HackerNews style forum can be built by composing a `comment' concept recursively to get threading and also composing with a `scoring' concept to get voting and sorting of comments~\cite{dejavu}.

The low-code syntax Mavo has been used to create a real-time messaging client using just modified HTML, although it is unclear whether the affordances offered by Mavo scale to creating a larger scale social application~\cite{mavo,mavochat}.
% This approach does not handle operation, although perhaps it would be possible to impliment these primitives in an interoperable way.
% However, as we discuss in Reification, we choose to make our

Tools like Hometown and Smalltown, built on Mastodon, and PubHubs, built on Matrix, make it easier for users to start up social sites for small communities by adding features like easier to use moderation tools and user interfaces aimed at non-technical users.
All of these projects add features that prioritize within-server interactions, and PubHubs disables federated interactions all together~\cite{smalltown,runyourownsocial,pubhubs,pubhubsidentity}.
This means that the communities created on top of these projects are bound to a particular user experience unless they migrate to a different instance where they risk losing some or all of their interactions with the community

The Initiative for Digital Public Infrastructure has proposed the creation of a ``Friendly Neighborhood Algorithm Store'' that would include bre-built solutions for detecting spam, mis/disinformation, harassment and CSAM as well as providing various ``For You'' recommendation algorithms.
The store has not yet been realized~\cite{threeleggedstool}.

\subsection{Adversarial Interoperability}

Adversarial interoperability is when one service connects with existing services without their permission~\cite{adverserialinterop}.
In the 1990's, the chat client ICQ connected MSN, Yahoo!, and AOL messaging services without their permission and it was so popular that the existing services gave in to its existence~\cite{threeleggedstool}.
Today, Matrix has built `bridges' between various closed messaging services including iMessage, Discord, Instagram, etc.

In the broader landscape of social media, the system Gobo integrates posts from Twitter, Mastodon, and Reddit all in one place. It then allows users to filter and sort this content with novel algorithms~\cite{gobo}.

Adversarial services necessarily cater to the lowest common denominator of different social applications designs.
While we believe these tools are certainly important for \emph{adoption} of new tools, they don't inherently broaden the design space of social applications, as we aim to in this work.

\subsection{Legally Enforced Interoperability}

Graffiti aims to make social applications easier to create by reducing development load and reducing lock-in to applications.
While we take a technical approach to tackle both problems simultaneously in a \emph{new} ecosystem, legal work is being done to deal with platform lock-in to the \emph{existing} ecosystem of social applications.
While a full survey of legal work is beyond the scope of this paper, we highlight two major works: the General Data Protection Regulation (GDPR) and the Digital Markets Act (DMA), both passed in the European Union.

\subsubsection{GDPR} Adopted in 2016 and effective since 2018, the GDPR provides a \emph{right of data portability}
to ``foster opportunities for innovation by means of sharing of personal data between data controllers.''
% These rights allow users to receive a copy of their `data provided' to a platform, however a narrow interpretation of this wording does not include data that a platform has \emph{observed} about a user.
However, while the legislation states that platforms ``should be encouraged to develop interoperable formats that enable data portability'' there is no explicit provision that they must, and while users ``should have the right to have the personal data transmitted directly from one controller to another,'' it is only where ``technically feasible''~\cite{gdpr,gdprinteroperability}.

In practice, this leads to an `adieu scenario' where users can remove their data from one platform and bring it another, often with substantial difficulty~\cite{howportable}. However some legal scholars are optimistic that it could be expanded to a `fusing scenario,' a ``user-centric platform where all digital services are interconnected''~\cite{gdprinteroperability}.
Perhaps Graffiti or a similar system could be required by law as the interoperable format and means of data transmission between platforms.

\subsubsection{DMA} Adopted in 2022, the DMA enforces that specific `gatekeepers,' including Meta, Apple, and ByteDance must make their \emph{messaging} services interoperable.
While the legislation does not specify what protocol gatekeepers are supposed to use to interoperate, it must allow for the interoperation of multimedia point-to-point and group texting as well as voice and video calls.
The gatekeepers have several years to comply, and as of writing, interoperability has not been implemented.
The legislation does not extend to include more general social media~\cite{digitalmarketsact}.

% hile we are in support of these systems, this sort of legislation will lag behind the technology, so we are attempting to envision a system that leads to better outcomes from the getgo.
% This paper makes the stand that there is a systematic problem with social media architecture that must be. However, there is plenty of work t
% While this paper takes a technosolutionist approach to a social problem.
% Digital Markets Act: interoperability between messaging applications.
% Some of these
% Issues with social media rile, from it's.
% There are attempts to curtail the problem with legal, li
% For example GDPR or a proposed ban on TikTok.
% We see this as treating the symptoms, not the underlying problem: if users had more


% \subsection{Three Legged Stool}

% The Initiative for Digital Publish Infrastructure has a "three-legged" proposal to fix social media: first they want to uplift usage of Very Small Online Platforms (VSOPs) while recognizing that VSOPs won't entirely replace massive-scale platforms; second they want to create "loyal clients" that aggregate content from different platforms, making it easier to belong to multiple platforms at the same time; third, they want to create a "Friendly Neighborhood Algorithms Store" full of pre-built recommendation algorithms, moderation systems and other tools that VSOPs can use to build up their own platforms with relative ease~\cite{threeleggedstool}.

% In many respects our work is an answer to this initiative: Graffiti makes it easy to create VSOPs and the univeral backend naturally connects them - there is no need for the adverserial interoperability~\cite{adverserialinterop} that the IDPI suggests is necessary to create loyal clients.

% While truly creating a ``Friendly Neighborhood Algorithm Store'' is beyond the scope of this, we imagine many tools built upon Graffiti including algorithms, moderation, as well as design frameworks - our Vue plugin for graffiti is a first step in that direction.
% The friendly neighborhood algorithm store itself could be implimented in Graffiti, which would mean that it wouldn't be able to lock users in unlike the Apple app store which has strict developer requirements.

% % We agree wholeheartedly with the first and third points:
% We too imagine that in the Graffiti ecosystem there are different scales of interaction, local communities, niches, as well as the massive scale networks like TikTok, X, or Facebook.
% By moving computation to client side, Graffiti reduces the complexity of building apps and our Vue plugin is just a start to the marketplace of development tools that could be built on top.
% With regard to the loyal client, The IDPI imagines these clients need to perform "adverserial interoperation", as done with the Gobo platform CITE - Graffiti may need to too, to acheive more widespread adoption.
% Moreover, aggregation will move beyond feed-based systems.

The At Protocol has flexible schemas but servers will only handle certain types of objects,
breaking our serverless requirement.
At proto is not serverless since it restricts the type of objects that can be placed on it.
While there may need to be some filtering for spam and illegal content, no strict set of schemas
The At Protocol, to our knowledge does not add such side effects and has invented its own vocabulary, similar to the Activity Vocabulary, and lexicon for extensible object definitions, similar to JSON Schema.
Our design is similar, starting from Activity Vocabulary rather than the At Protocol's syntax, but we do not \emph{require} applications to specify schemas for the data they produce. Doing so would put up more barriers to less-technical users; the semantic web's slow adoption has made it clear that users don't like to add metadata~\cite{semanticwebtwodecadeson}.
% The Activity Vocabulary is an essential step in this direction that provides an extensive and extensible vocabulary for objects and activities.
% However, the ActivityPub protocol that builds on  the vocabulary undercuts the reification benefits by adding \emph{side effects} to the creation or reception of various activities and object, that undoes the reification.
% As mentioned, these side effected are implemented inconsistently across servers and changing or adding new functionality requires changing servers, violating our server-less design
\subsubsection{Comparison to Other models}

While channels are a relatively simple, they are subtly differences from other established concepts:

The channel model is clearly different from the Firehose Model used by BlueSky
which.
The firehose model dumps all user data into one public database,
which doesn't allow for the carving out of different contexts that we did in our comment
example above. In the firehose model a comment will always be visible to *both* the original post's audience and
the commenter's followers.
It also has subtle differences.
There are several other contextual models.

\paragraph{Actor}

We use actors to describe identities in the system,
but ActivityPub uses the actor model to describe the system itself.
As mentioned the actor model used by ActivityPub is a fusion of
direct messaging (like Email) and broadcasting like RSS.
Having channels

Compared to the Actor Model it removes the need for Server-Side effects
The actor model is a fusion of direct messaging (like Email) and broadcasting
(like RSS) and works well for follow-based communication but struggles
to pass information via other rendez-vous.
In the actor model, even something as simple as comments can be
[very tricky and require server "side effects"](https://seb.jambor.dev/posts/understanding-activitypub-part-3-the-state-of-mastodon/).

\paragraph{Bidirectional Links}

On the World Wide Web, hyperlinks point in one direction from one website to another.
However earlier hypermedia systems, like Xanadu and Dexter, included \emph{bidirectional} links:
from one document you could see all the other documents that link to it.
Bidirecitonal links are now currently possible in personal information management systems like Roam and Notion.

Channels can be seen as a way to \emph{selectively} create bidirectional links.
A reply object will have pointers to the object it is replying to and the actor who created the reply.
Adding the object's URI or the actor's URI as channels selectively adds backlinks to those entities
making it possible to go \emph{from} the original post \emph{to} the reply (in the case of Instagram-like comments)
or \emph{from} the actor \emph{to} the reply (in the case of Twitter/X-like comments).
Of course, channels also make it possible to create "links" from non-digital artifacts like topics or locations.

\paragraph{Object Capabilities}

Object capability security (OCaps) is a security model where rights to perform an action
can be transferred to another user by giving that user a reference to that action.
Channels can be thought of as a capability mechanism where
capability read or write to a particular channel
can be transferred to another user by telling them the channel's string.

However Graffiti isn't a purely capability-based system because it
is also possible to add an "allow" list to each object.
For example a group chat could use a private string as a channel
and if that string is sufficiently long and random so as not to be guessed
the conversation would remain private.
However this makes it possible for any member of the group to share (or accidentally leak) the
channel and let anyone else in.
Allow lists provide more long term privacy.

Allow lists are managed by applications.
If allow lists are especially enourmous then is it really public?
